---
title: "Hotel_Booking_Analysis"
author: "Mohd Javed"
date: "`r Sys.Date()`"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,cache = TRUE,warning = FALSE,message = FALSE,dpi = 180,fig.width = 8,fig.height = 5)
```


### Importing required pacakges


```{r}
library(readr)
library(magrittr)
library(dplyr)
library(ggplot2)
library(GGally)
library(skimr)

```


```{r}
### importing dataset


hotels <- read_csv("hotels.csv")
#View(hotels)
```


### 


```{r}

hotel_stays<-hotels %>%
  filter(is_canceled==0)%>%
  mutate(children=case_when(children+babies>0 ~ "children",TRUE~"none"),required_car_parking_spaces=case_when(required_car_parking_spaces>0~"parking",TRUE~"none"))%>%
  select(-is_canceled,-reservation_status,-babies)

hotel_stays%>%
  count(children)
```


### Exploratory Data Analysis


```{r}
skim(hotel_stays)
```


### Cancellation proportion by Arrival date month  


```{r}
hotel_stays%>%
  mutate(arrival_date_month=factor(arrival_date_month,levels = month.name))%>%
  count(hotel,arrival_date_month,children)%>%
  group_by(hotel,children)%>%
  mutate(proportion=n/sum(n))%>%
  ggplot(aes(arrival_date_month,proportion,fill=children))+
  geom_col(position = "dodge")+
  scale_y_continuous(labels = scales::percent_format())+
  facet_wrap(~hotel,nrow = 2)

```



### Hotel type proportion by required parking spaces 



```{r}
hotel_stays%>%
  count(hotel,required_car_parking_spaces,children)%>%
  group_by(hotel,children)%>%
  mutate(proportion=n/sum(n))%>%
  ggplot(aes(required_car_parking_spaces,proportion,fill=children))+
  geom_col(position = "dodge")+
  scale_y_continuous(labels = scales::percent_format())+
  facet_wrap(~hotel,nrow = 2)


```


### Multiple plots of children, adr, car_parking_spaces and total number of special requests




```{r}
hotel_stays%>%
  select(children,adr,required_car_parking_spaces,total_of_special_requests)%>%
  ggpairs(mapping = aes(color=children))
```


### filtering variables for model





```{r}
hotel_DF<-hotel_stays%>%
  select(children,hotel,arrival_date_month,meal,adr,adults,required_car_parking_spaces,total_of_special_requests,stays_in_week_nights,stays_in_weekend_nights)%>%
  mutate_if(is.character,factor)
skim(hotel_DF)

```


### Data wrangling 



```{r}
# importing required packages for model training
library(tidymodels)
library(recipes)
library(themis)

set.seed(123)
hotel_split <- initial_split(hotel_DF)
hotel_train_set <- training(hotel_split)
hotel_test_set <- testing(hotel_split)
hotel_recipe <- recipe(children ~ ., data = hotel_train_set) %>%
  step_downsample(children) %>%
  step_dummy(all_nominal(), -all_outcomes()) %>%
  step_zv(all_numeric()) %>%
  step_normalize(all_numeric()) %>%
  prep()
hotel_recipe


```


### Model training


```{r}
test_proce <- bake(hotel_recipe, new_data = hotel_test_set)
juice(hotel_recipe) %>% count(children)
```



### K nearest neighbour 


```{r}
knn_spec<-nearest_neighbor()%>%
  set_engine("kknn")%>%
  set_mode("classification")
library(kknn)
knn_fit<-knn_spec %>%
  fit(children~.,
      data=juice(hotel_recipe))




```


### Training Decision Tree (recursive partitioing)



```{r}
tree_spec<-decision_tree()%>%
  set_engine("rpart")%>%
  set_mode("classification")
tree_fit<-tree_spec %>%
  fit(children~.,data=juice(hotel_recipe))
tree_fit
```





```{r}
validation_splits <-mc_cv(juice(hotel_recipe), prop = 0.9, strata = children)
```


#### Training KNN model


```{r}
knn_res <- fit_resamples(
  knn_spec,
  children ~ ., 
  validation_splits,
  control = control_resamples(save_pred = TRUE)
)
        
knn_res %>%
  collect_metrics()
```



### Validation using validation splits


```{r}
tree_res <- fit_resamples(tree_spec,
                          children ~ .,
                          validation_splits,
                          control = control_resamples(save_pred = TRUE))

tree_res %>%
  collect_metrics()
```


```{r}

knn_res%>%
  unnest(.predictions)%>%
  mutate(model="kknn")%>%
  bind_rows(tree_res%>%
  unnest(.predictions)%>%
  mutate(model="rpart"))%>%
  group_by(model)%>%
  roc_curve(children,.pred_children)

```

